---
title: "Reproducible Research: Peer Assessment 1"
output: 
 html_document: 
    keep_md: true
---


```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```


## Loading and preprocessing the data
# 1.Load the data (i.e. read.csv())
# 2. 2.Process/transform the data (if necessary) into a format suitable for your analysis
unzip the data file and then read data file
```{r Loading and preprocessing the data, results=FALSE}
unzip("activity.zip", file=NULL, list=FALSE, overwrite=TRUE)  
REPW1Data <- read.csv("activity.csv")
library(dplyr)
REPW1Data_df <- tbl_df(REPW1Data)
REPW1Data_df_group_By_Date <- group_by(REPW1Data_df, date)
REPW1Data_df_Sum_By_Date <- summarize(REPW1Data_df_group_By_Date, steps=sum(steps))
```

  
## What is mean total number of steps taken per day?
#1.Calculate the total number of steps taken per day
#2.If you do not understand the difference between a histogram and a barplot, research the difference between them. Make a histogram of the total number of steps taken each day
```{r hist, fig.height=4}
hist(REPW1Data_df_Sum_By_Date$steps, col="gray", ylab="Days", xlab="Steps", main="Total Number Of Steps Taken Per Day")
rug(REPW1Data_df_Sum_By_Date$steps)
```

#3.Calculate and report the mean and median of the total number of steps taken per day
```{r mean and median total}
mean(REPW1Data_df_Sum_By_Date$steps, na.rm=TRUE)
median(REPW1Data_df_Sum_By_Date$steps, na.rm=TRUE)
```


## What is the average daily activity pattern?
#1.Make a time series plot (i.e. type = "l") of the 5-minute interval (x-axis) and the average number of steps taken, averaged across all days (y-axis)
```{r average daily activity pattern}
library(ggplot2)
REPW1Data_df_group_By_Time <- group_by(REPW1Data_df, interval)
REPW1Data_df_Mean_By_Time <- summarize(REPW1Data_df_group_By_Time, steps=mean(steps,na.rm=TRUE))
ggplot(REPW1Data_df_Mean_By_Time, aes(interval, steps)) + geom_line() + labs(title = "Time series plot of the average number of steps taken", x="Time Series", y="Average Number Of Stpes")
```

#2.The 5-minute interval that, on average, contains the maximum number of steps 
```{r}
REPW1Data_df_Mean_By_Time[which.max(REPW1Data_df_Mean_By_Time$steps),]    
```

## Imputing missing values
# 1.Calculate and report the total number of missing values in the dataset
```{r Calculate total missing values}
    nrow(REPW1Data_df[is.na(REPW1Data_df$steps),])
```
# 2.Devise a strategy for filling in all of the missing values in the dataset.
  replace to NA value with mean steps in that time
  
# 3.Create a new dataset that is equal to the original dataset but with the missing data filled in.
```{r imputation missing values}
NewImputedTbl <- data.frame()
for(i in 1:nrow(REPW1Data_df)){
    valSteps <- REPW1Data_df[i,]
    if(is.na(valSteps$steps)) {
        NewImputedTbl[i,c("steps")] <-subset(REPW1Data_df_Mean_By_Time,interval==valSteps$interval, select="steps")}
    else{
        NewImputedTbl[i, c("steps")] <- valSteps[,"steps"]
        }
}

REPW1Data_df_Imputted <- REPW1Data_df
REPW1Data_df_Imputted$steps <-NewImputedTbl$steps
```

# 4.Make a histogram of the total number of steps taken each day and Calculate and report the mean and median total number of steps taken per day.
```{r original & imputted data histogram}
par(mfrow=c(2,1))
# Create Total number of Steps Histogram(Original)
REPW1Data_df_Sum_By_Date <- summarize(REPW1Data_df_group_By_Date, steps=sum(steps))
hist(REPW1Data_df_Sum_By_Date$steps, col="gray", ylab="Days", xlab="Steps", main="Mean Number Of Steps Taken Per Day")
rug(REPW1Data_df_Sum_By_Date$steps)

# Create Total number of Steps Histogram(Imputing missing values)
REPW1Data_df_Imputted_group_By_Date <- group_by(REPW1Data_df_Imputted, date)
REPW1Data_df_Imputted_Sum_By_Date <- summarize(REPW1Data_df_Imputted_group_By_Date, steps=sum(steps))
hist(REPW1Data_df_Imputted_Sum_By_Date$steps, col="gray", ylab="Days", xlab="Steps", main="Mean Number Of Steps Taken Per Day(Imputing missing values)")
rug(REPW1Data_df_Imputted_Sum_By_Date$steps)
```

# Mean and Median value of the imputted data
```{r compae mean and median number}
mean(REPW1Data_df_Imputted_Sum_By_Date$steps, na.rm=TRUE)
median(REPW1Data_df_Imputted_Sum_By_Date$steps, na.rm=TRUE)
```


## Are there differences in activity patterns between weekdays and weekends?
